EGGs are adhesive!

### 2025-03-11 calco/main.tex
cose

We have now all the ingredient necessary to study the adhesivity properties of $\EqHyp$.  As a first step we need to introduce a class $\mathcal{M}$ of monos.

\begin{definition}
	contenuto...
\end{definition}

\begin{restatable}{lemma}{pbm}\label{lem:pbmono}
	The class $\pbc$ is closed under composition, decomposition and it is stable under pullbacks and pushouts
\end{restatable}

@@

\begin{lemma}\label{lemma:stab}
	In $\EqHyp$, $\reg(\EqHyp), \mathcal{N}$-pushouts are stable.
\end{lemma}

@@

We have to show that the top face of the cube with which we have began is a pushout. Let $\mathcal{H}$ be $(E, V, C, s, t, q)$ and suppose that there exists  $o\colon  \mathcal{G}_2' \to \mathcal{H}$ and $w\colon  \mathcal{G}_3' \to \mathcal{H}$ such that $o \circ h' = w \circ k'$. Since $T$ preserves colimits by \Cref{prop:forghyp}, we know that there exists a morphism $(v_E, v_V)\colon T(\mathcal{G'}_4)\to T(\mathcal{H})$ such that	
\[v_E\circ h'_E=w_E\circ k'_E\qquad v_V\circ h'_V=w_V\circ k'_V\]

We want to extend such a morphism to one of $\EqHyp$ between $\mathcal{G}'_4\to \mathcal{H}$. If we are able to do so we can conclude because uniqueness is guaranteed by \Cref{rem:eqhyp_morphs}.

Now, $d_V$ is a regular epi, thus

@@

	We want to construct the dotted arrow $v_C$ which fits in the diagram below.
	\[
	\xymatrix@C=10pt@R=6pt{
		&B_1'\ar[dd]|\hole_(.65){q_1'}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]|\hole_(.65){q_2'} \ar[dl]_{t_V'}\ar[dr]^{z_V}\\
		B_3'  \ar[dd]_{q_3'}\ar[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){q_4'} \ar[rr]^(.7){v_V}&&  V\ar[dd]^{q}\\
		&C_1'\ar[rr]|\hole^(.65){h_C'} \ar[dl]_{k_C'} && C_2' \ar[dl]^{t_C'}\ar[dr]^{z_C}\\
		C'_3 \ar[rr]_{p_C'} & & C'_4\ar@{.>}[rr]_{v_C} && C
	}\]
	By , we know that the top face of the cube below is a pushout.
	\[\xymatrix@C=15pt@R=15pt{&K_{s_2\circ m_1\circ q'_1}\ar[dd]|\hole_(.65){\pi_{s_2\circ m_1\circ q'_1}^{1}}\ar[rr]^{k_{h'_2}} \ar[dl]_{k_{k'_2}} && K_{m_2\circ q'_2} \ar[dd]^{\pi_{m_2\circ q'_2}^{1}} \ar[dl]_{k_{t'_2}} \\ K_{m_3\circ q'_3}  \ar[dd]_{\pi_{m_3\circ q'_3}^{1}}\ar[rr]^(.7){k_{p'_2}} & & K_{q'_4} \ar[dd]_(.3){\pi_{q'_4}^{1}}\\&B'_1\ar[rr]|\hole^(.65){h'_2} \ar[dl]^{k'_2} && B'_2 \ar[dl]^{t'_2} \\B'_3 \ar[rr]_{p'_2} & & C_4}\]
	And, since $m_3$ and $m_2$ are monos,
	\[\begin{split}
		q'_3 \circ \pi_{m_3 \circ q'_3}^1 = q'_3 \circ \pi_{m_3\circ q_3'}^2
	\end{split}\qquad\begin{split}
		q_2' \circ \pi_{m_2 \circ q_2'}^1 = q'_2 \circ \pi_{m_2 \circ q_2'}^2
	\end{split}\]
	Computing, we obtain
	\[\begin{split}
		q \circ v_V \circ \pi_{q_4'}^1 \circ k_{p_V'} &= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^1 \\
		&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^1 \\
		&= w_C \circ q_3' \circ \pi_{m_3 \circ q_3'}^1 \\
		&= w_C \circ q_3' \circ \pi_{m_3 \circ q_3}^2 \\
		&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^2 \\
		&= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^2 \\
		&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{p'_V}
	\end{split}\qquad\begin{split}
		q \circ v_V \circ \pi_{q'_4}^1 \circ k_{t_V'} &= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^1 \\
		&= q \circ z_V \circ \pi_{m_2 \circ q_2'}^1 \\
		&= z_C \circ q_2' \circ \pi_{m_2 \circ q_2'}^1 \\
		&= z_C \circ q_2' \circ \pi_{m_2 \circ q_2}^2 \\
		&= q \circ z_V \circ \pi_{m_2 \circ q_2'}^2 \\
		&= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^2 \\
		&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{t'_V}
	\end{split}\]
	Since the previous cube has a pushout as top face, by universal property, we have
	\[
	q \circ v_V \circ \pi_{q'_4}^1 = q \circ v_V \circ \pi_{q'_4}^2
	\]
	hence, $v_C$ is the mediating arrow.
	\[
	v_C \circ q'_4 \circ \pi_{q'_4}^1 = v_C \circ q'_4 \circ \pi_{q'_4}^2
	\]

%	Let now $z: \graph{G}'_2 \to \graph H = (E, V, Q, s, t, q)$, $w: \graph G_3^1 \to \graph H$ be two morphisms such that $z \circ h' = w \circ k'$, and let $\phi:B'_4 \to V$ be the arrow induced by $z_V$ and $w_V$. We want to construct the arrow $v: C'_4 \to Q$ in the diagram below.
	\[\begin{tikzcd}[row sep = 20, column sep = 20]
		& {B_1'} && {B_2'} \\
		{B_3'} && {B_4'} && V \\
		& T && U \\
		Y && {C_4'} && Q \\
		& {C_1} && {C_2} \\
		{C_3} && {C_4}
		\arrow["{h_V'}", from=1-2, to=1-4]
		\arrow["{k_V'}"', from=1-2, to=2-1]
		\arrow["{s_2\circ m_1 \circ q_1'}"{description, pos=0.7}, from=1-2, to=3-2]
		\arrow["{t_V'}", from=1-4, to=2-3]
		\arrow["{z_V}", from=1-4, to=2-5]
		\arrow["{m_2\circ q'_2}"{pos=0.7}, from=1-4, to=3-4]
		\arrow["{m_3 \circ q'_3}"', from=2-1, to=4-1]
		\arrow["q", from=2-5, to=4-5]
		\arrow["{x_1}"{pos=0.8}, from=3-2, to=3-4]
		\arrow["w"', from=3-2, to=4-1]
		\arrow["{x_2}"'{pos=0.7}, from=3-2, to=5-2]
		\arrow["{u_1}", from=3-4, to=4-3]
		\arrow["{u_2}"{pos=0.7}, from=3-4, to=5-4]
		\arrow["{y_2}"', from=4-1, to=6-1]
		\arrow["{h_C}"{pos=0.8}, from=5-2, to=5-4]
		\arrow["{k_C}", from=5-2, to=6-1]
		\arrow["{t_C}", from=5-4, to=6-3]
		\arrow["{p_C}"', from=6-1, to=6-3]
		\arrow["{p_V'}"{pos=0.8}, from=2-1, to=2-3, crossing over]
		\arrow["\phi"{pos=0.8}, from=2-3, to=2-5, crossing over]
		\arrow["{q'_4}"'{pos=0.7}, from=2-3, to=4-3, crossing over]
		\arrow["{y_1}"{pos=0.7}, from=4-1, to=4-3, crossing over]
		\arrow["{d_C}"'{pos=0.7}, from=4-3, to=6-3, crossing over]
		\arrow["v"{pos=0.8}, dashed, from=4-3, to=4-5, crossing over]
	\end{tikzcd}\]

@@

 We are left with uniqueness. If $\theta'$ is another arrow $K_d\to Z$ such that $\theta'\circ k_{n'}=\gamma$ and $\theta'\circ k'_{g'}= \delta$. If we compute we get
 \[\theta'\circ j_3=\theta'\circ k_{g'}\circ x_3=\delta \circ x_3=\theta \circ k_{g'}\circ x_3=\theta \circ j_3 \]
 We already now that $\theta'\circ k_{n'}=\theta \circ k_{n'}$, so the previous identity entails that $\theta =\theta'$ as wanted.
			 \qedhere 
	\end{enumerate}
\end{proof}

>>>>

We have now all the ingredient necessary to study the adhesivity properties of $\EqHyp$.  As a first step we need to introduce a class of monos.

\begin{definition}
	We define $\pbc$ as the class of all regular monos $(h_E, h_V, h_Q)\colon \mathcal{G}\to \mathcal{H}$ in $\EqHyp$ such that the square on the right is a pullback
\end{definition}

\begin{restatable}{lemma}{pbm}\label{lem:pbmono}
	The class $\pbc$ is closed under composition, decomposition and it is stable under pullbacks and pushouts.
\end{restatable}
\begin{proof}
	Closure under composition and decomposition follows from \Cref{lem:pb1}, while stability under pushouts is the first point of \Cref{prop:kerset}. We are left with stability under pullbacks.
\end{proof}

@@

\begin{lemma}\label{lemma:stab}
	In $\EqHyp$, $\pbc$-pushouts are stable.
\end{lemma}

@@

We have to show that the top face of the cube with which we have began is a pushout. Let $\mathcal{H}$ be $(E, V, Q, s, t, q)$ and suppose that there exists  $o\colon  \mathcal{G}_2' \to \mathcal{H}$ and $w\colon  \mathcal{G}_3' \to \mathcal{H}$ such that $o \circ h' = w \circ k'$. Since $T$ preserves colimits by \Cref{prop:forghyp}, we know that there exists a morphism $(v_E, v_V)\colon T(\mathcal{G'}_4)\to T(\mathcal{H})$ such that	$v_E\circ h'_E=w_E\circ k'_E$ and $v_V\circ h'_V=w_V\circ k'_V$.

\noindent 
\parbox{9cm}{ We want to extend such a morphism to one of $\EqHyp$ between $\mathcal{G}'_4\to \mathcal{H}$. If we are able to do so we can conclude because uniqueness is guaranteed by \Cref{rem:eqhyp_morphs}.  Concretely we have to construct the dotted arrow $v_C$ which fits in the diagram aside.} \hfill \parbox{3cm}{
\xymatrix@C=10pt@R=6pt{
	&B_1'\ar@{>>}[dd]|\hole_(.65){q_1'}\ar@{>->}[rr]^{h_V'} \ar[dl]_(.6){k_V'} && B_2' \ar@{>>}[dd]|\hole_(.65){q_2'} \ar[dl]_(.6){z_V'}\ar[dr]^{o_V}\\
	B_3'  \ar@{>>}[dd]_{q_3'}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar@{>>}[dd]_(.3){q_4'} \ar[rr]^(.7){v_V}&&  V\ar[dd]^{q}\\
	&Q_1'\ar@{>->}[rr]|\hole^(.65){h'_Q} \ar[dl]_{k'_Q} && Q_2' \ar[dl]^(.4){z_Q'}\ar[dr]^{o_Q}\\
	Q'_3 \ar@{>->}[rr]_{p_Q'} & & Q'_4\ar@{.>}[rr]_{v_C} && Q
}}

Now, consider the cube below, $h$ is in $\pbc$ so by \Cref{lem:pb1} we know that its back face is a pullback. We can then apply the second point of \Cref{prop:kerset} to deduce that the square on the right is a pushout.
\[\xymatrix@C=10pt@R=10pt{
	&B_1'\ar@{>>}[dd]|(.53)\hole^(.67){s_2\circ m_1\circ q_1'}\ar@{>->}[rrrr]^{h_V'} \ar[dl]_(.6){k_V'} &&&& B_2' \ar@{>}[dd]^{m_2\circ q_2'} \ar[dl]_(.6){z_V'} & & K_{s_2\circ m_1\circ q'_1}  \ar@{>->}[rr]^-{k_{x_1}} \ar[dd]_{k_{w}}&& K_{m_2\circ q'_2} \ar[dd]^{k_{u_1}}\\
	B_3'  \ar[dd]_{m_3\circ q_3'}\ar@{>->}[rrrr]^(.7){p_V'} & & &&B_4' \ar@{>>}[dd]_(.3){q_4'} \\
	&T\ar@{>->}[rrrr]|(.68)\hole_{x_1} \ar[dl]_{w} &&&& U\ar[dl]^(.4){u_1} && K_{m_3\circ q'_3} \ar@{>->}[rr]_-{k_{y_1}} && K_{q'_4} \\
	Y \ar@{>->}[rrrr]_{y_1} & & && Q'_4}\]

By construction we have that 	$m_3\circ q'_3 \circ \pi_{m_3 \circ q'_3}^1 = m_3\circ q'_3 \circ \pi_{m_3\circ q_3'}^2$ and

Since $m_3$ and $m_2$ are monos we have
\[\begin{split}
	q'_3 \circ \pi_{m_3 \circ q'_3}^1 = q'_3 \circ \pi_{m_3\circ q_3'}^2
\end{split}\qquad\begin{split}
	q_2' \circ \pi_{m_2 \circ q_2'}^1 = q'_2 \circ \pi_{m_2 \circ q_2'}^2
\end{split}\]
Computing, we obtain
\[\begin{split}
	q \circ v_V \circ \pi_{q_4'}^1 \circ k_{p_V'} &= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^1 \\
	&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^1 \\
	&= w_C \circ q_3' \circ \pi_{m_3 \circ q_3'}^1 \\
	&= w_C \circ q_3' \circ \pi_{m_3 \circ q_3}^2 \\
	&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^2 \\
	&= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^2 \\
	&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{p'_V}
\end{split}\qquad\begin{split}
	q \circ v_V \circ \pi_{q'_4}^1 \circ k_{t_V'} &= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^1 \\
	&= q \circ z_V \circ \pi_{m_2 \circ q_2'}^1 \\
	&= z_C \circ q_2' \circ \pi_{m_2 \circ q_2'}^1 \\
	&= z_C \circ q_2' \circ \pi_{m_2 \circ q_2}^2 \\
	&= q \circ z_V \circ \pi_{m_2 \circ q_2'}^2 \\
	&= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^2 \\
	&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{t'_V}
\end{split}\]
Since the previous cube has a pushout as top face, by universal property, we have
\[
q \circ v_V \circ \pi_{q'_4}^1 = q \circ v_V \circ \pi_{q'_4}^2
\]
hence, $v_C$ is the mediating arrow.
\[
v_C \circ q'_4 \circ \pi_{q'_4}^1 = v_C \circ q'_4 \circ \pi_{q'_4}^2
\]

@@

@@

 We are left with uniqueness. If $\theta'$ is another arrow $K_d\to Z$ such that $\theta'\circ k_{n'}=\gamma$ and $\theta'\circ k'_{g'}= \delta$. If we compute we get
 \[\theta'\circ j_3=\theta'\circ k_{g'}\circ x_3=\delta \circ x_3=\theta \circ k_{g'}\circ x_3=\theta \circ j_3 \]
 We already know that $\theta'\circ k_{n'}=\theta \circ k_{n'}$, so the previous identity entails that $\theta =\theta'$.
			 \qedhere 
	\end{enumerate}
\end{proof}
